OpenMPIR

### 2017-09-07 paper.tex
Addressing Billy's concerns re: performance conclusions

While we would like to distinguish between these, it is difficult to do so with
only performance counter information. A more thorough investigation would
require careful knowledge of each of the frontends and runtimes. One thing we
can guess at is that despite the focus of the paper, \emph{IR optimisations are
unlikely to be the cause of performance improvements}. Our evidence for this
comes from Shardl et al. \cite{tapir}. While some IR optimisations have been
implemented, they seem to have significantly more effect on parallel loop code,
while the tasking code remains relatively unaffected by IR optimizations
\cite{tapir}.

@@

This confirms our suspicions that most of the slowdown incurred is overhead in
the runtime. For \texttt{tapir}, the overhead is relatively small given the
size of the work chunks, for \texttt{clang} and \texttt{icc} it is
significantly larger, while for \texttt{gcc} it is completely overwhelming.
This does let us rule out the possibility that somehow the only the codegen for
other implementations is the culprit: it is almost certainly at least partially
the runtime implementation.

@@

The goal of this work is to show that a single intermediate representation can be
used to implement multiple programming models, with good performance. Schardl et 
al. showed that Tapir works well as an IR for Cilk, and surmised that it could
be an effective IR for a large subset of OpenMP. This work has taken the first
step in that direction, showing that it can work well as an IR for OpenMP tasks.

@@

>>>>

While we would like to distinguish between these, it is difficult to do so with
only performance counter information. A more thorough investigation would
require deep knowledge of each of the frontends and runtimes. While the
importance of optimizations of parallel code is clear, we hypothesize that most
of the disparity present in these results is a function of runtime
implementation differences. Our evidence for this comes from Shardl et al.
\cite{tapir}. Generally, the effect of the optimizations that have been
implemented for Tapir IR are outweighed by runtime contention, memory accesses,
etc. for parallel runs. \gsnote{Billy, this OK? Don't want to sound too
negative about the importance of the optimizations you've implemented.}

@@

This supports our hypothesis that most of the slowdown incurred is overhead in
the runtime. For \texttt{tapir}, the overhead is relatively small given the
size of the work chunks, for \texttt{clang} and \texttt{icc} it is
significantly larger, while for \texttt{gcc} it is completely overwhelming.

@@

The goal of this work is to show that a single intermediate representation can be
used to implement multiple programming models, with good performance. Schardl et 
al. showed that Tapir works well as an IR for Cilk, and surmised that it could
be an effective IR for a large subset of OpenMP \cite{tapir}. This work has
taken the first step in that direction, showing that it can work well as an IR
for OpenMP tasks.

@@

An obvious first step towards better OpenMP coverage, other than the features
listed above, is OpenMP \texttt{parallel for} loops. These are similar enough
to \texttt{cilk\_for} loops in semantics that this also should be a relatively
painless next step to adapt from the existing Cilk implementation.
